Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-08-17 doc/report/msc-report.tex
includegraphics - needs size.
little fixes.
report: fixed alignment of Zohmg diagrams.
report: de-indeed-ification

BigTable's data model can indeed be thought of as a sparse, distributed, sorted,
multidimensional map. In concept, this multidimensional map is identical to a
nested hash table. The dimensions of the map are: row key, column family and
column qualifier, and a version time-stamp.

@@

As the name implies, MapReduce is made up of two parts or phases: map and
reduce. The user specifies a map function that consumes a single key/value pair
to generate a set of intermediate key/value pairs, and a reduce function that
merges all intermediate values that belongs to the same key. The abstraction of
two functions to process small parts of the large data set was inspired by the
map and reduce functions commonly found in functional programming languages such
as Lisp. The use of this model allows the framework to easily parallelize the
computations. In the inevitable case of a failure at a computational node, the
computations that failed can simply be re-executed on another node.

@@

@@

While relational databases can be made highly available and reliable with load
balancing and so called \textit{sharding}, the setup is non-trivial. Regarding
scalability, relational databases are highly dependent on having data in working
memory; thus they scale poorly.
% TODO: [citation needed]

@@

@@

@@

The reasons behind using Dumbo and writing Zohmg in Python are similar. Last.fm
was already moving towards writing a larger set of its analysis tools in Python
thanks to Dumbo, which is a product of another intern at the company. Python is
a language which allows for rapid prototyping and is as such well suited for
writing both the user's map function and indeed the whole of Zohmg.

The main point of contention was which backing store to use. A number of
different systems that fitted the requirements were tried out but none of them
proved to be either very dissimilar from the BigTable apporach of HBase, or
indeed ready for production. HBase has the advantage of being closely tied to
Hadoop which means that there was already tools available for writing
HBase-centric MapReduce jobs in Hadoop. The technology culture at Last.fm again
proved to be the main reason: there were already projects within the company
that were using HBase, so the needed infrastructure and know-how was there.

>>>>

BigTable's data model can be thought of as a sparse, distributed, sorted,
multidimensional map. In concept, this multidimensional map is identical to a
nested hash table. The dimensions of the map are: row key, column family and
column qualifier, and a version time-stamp.

@@

As the name implies, MapReduce is made up of two parts or phases: map and
reduce. The user specifies a map function that consumes a single key/value pair
to generate a set of intermediate key/value pairs, and a reduce function that
merges all intermediate values that belongs to the same key.

MapReduce is not a new concept. The abstraction of two functions to process
small parts of the large data set was inspired by the map and reduce functions
commonly found in functional programming languages such as Lisp.

@@

The use of this model allows the framework to easily parallelize the
computations. In the inevitable case of a failure at a computational node, the
computations that failed can simply be re-executed on another node.

@@

While relational databases can be made highly available and reliable with load
balancing and so called \textit{sharding}, the setup is non-trivial.

@@

\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.40]{zohmg-import.eps}
        \caption{Diagram over data import to Zohmg.}
    \end{center}
\end{figure}

@@

\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.35]{zohmg-export.eps}
        \caption{Diagram over data export from Zohmg.}
    \end{center}
\end{figure}

@@

The reasons behind using Dumbo and writing Zohmg in Python are similar. Last.fm
was already moving towards writing a larger set of its analysis tools in Python
thanks to Dumbo, which is a product of another intern at the company. Python is
a language which allows for rapid prototyping and is as such well suited for
writing both the user's map function and the whole of Zohmg.

The main point of contention was which backing store to use. A number of
different systems that fitted the requirements were tried out but none of them
proved to be either very dissimilar from the BigTable apporach of HBase, or
ready for production. HBase has the advantage of being closely tied to
Hadoop which means that there was already tools available for writing
HBase-centric MapReduce jobs in Hadoop. The technology culture at Last.fm again
proved to be the main reason: there were already projects within the company
that were using HBase, so the needed infrastructure and know-how was there.
